---
title: "Practicing R with the DataScienceExercises Package"
description: "How to install and use the DataScienceExercises interactive exercise package to practice R with immediate feedback."
author: "Claudius Gräbner-Radkowitsch"
date: 2026-06-16
date-modified: 2026-06-16
categories:
  - R
  - exercises
---

When learning a programming language, applying new concepts regularly is essential. Without regular practice it is hard to remember — and to find out what you have not yet understood well enough.

The `DataScienceExercises` package provides a set of interactive exercises for core R topics. Each exercise runs in your browser and gives immediate feedback on your answers. This post explains how to install the package and run exercises.

## Installation

The exercises use three packages: `learnr`, `gradethis`, and `DataScienceExercises`. All three are installed from GitHub, so you first need the `remotes` package:

```{r, eval=FALSE}
install.packages("remotes")
```

Then install the exercise packages:

```{r, eval=FALSE}
remotes::install_github(
  repo = c(
    "rstudio/learnr",
    "rstudio/gradethis",
    "graebnerc/DataScienceExercises"
  ),
  upgrade = "always"
)
```

## Keeping the package up to date

The exercises are updated and extended over time. Before starting a practice session, run:

```{r, eval=FALSE}
remotes::install_github(
  repo = "graebnerc/DataScienceExercises",
  upgrade = "always"
)
```

An internet connection is only needed to update — the exercises themselves run offline once installed.

## Running an exercise

Resource pages on this hub that have associated exercises list the name of the exercise set. To launch one, pass that name to `learnr::run_tutorial()`. For example, to run the `Basics` set:

```{r, eval=FALSE}
learnr::run_tutorial(
  name = "Basics",
  package = "DataScienceExercises",
  shiny_args = list("launch.browser" = TRUE)
)
```

Replace `"Basics"` with the name of whichever exercise set you want to practice.

## Reporting bugs or suggestions

If you find a mistake or have an idea for a new exercise, use the [issue tracker](https://github.com/graebnerc/DataScienceExercises/issues) on GitHub.
